### 2016-09-28 paper.tex
changes

%\newcommand{\smin}[1]{s_\text{min}({#1})}
\newcommand{\smin}{s_\text{min}}
\newcommand{\smax}{s_\text{max}}
\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Zproj}{Z^{\textit{proj}}}
\newcommand{\Zreopt}{Z^{\textit{owa}}}
\newcommand{\nreopt}{n^{\textit{owa}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\matW}{\hat W}
\newcommand{\matV}{\hat V}
\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Wowa}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\ahat}{\hat\alpha}
\newcommand{\afull}{\ahat^{\textit{full}}}
\newcommand{\wowa}{\hat\w^{owa}}
\newcommand{\wowafull}{\hat\w^{\textit{owa,full}}}
\newcommand{\wowastar}{\hat\w^{\textit{owa,*}}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

In this paper, we propose a \emph{semi-interactive} distributed algorithm called \emph{optimal weighted averaging} (OWA).
Our algorithm performs two rounds of communication,
so it is not subject to the existing regret bounds of non-interactive algorithms.
The amount of information communicated in the second round is small, however,
so our algorithm retains the speed advantages of non-interactive algorithms.
We now formally describe our problem and proposed solution.

\subsection{Problem Statement}
%\section{Problem Setting and Our Solution}

@@

%That is, we assume there exists a constant $c$ such that
%\begin{equation}
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
%.
%\end{equation}
\begin{assumption}
Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
\begin{equation}
%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
,
\end{equation}
where $\I_\wstar$ is the positive definite Fisher information matrix.
\end{assumption}

@@

Our condition is more general than the conditions required in previous work.
\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded.
This condition implies Condition 1 because every bounded random variable is automatically sub-Gaussian by definition.

@@

\begin{theorem}
\label{thm:wave}
Assume Condition 1 holds.
Let $t>0$.
Then with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
\end{equation}
where
\begin{equation}
v_t =
\sigma^2
\left(
\tr{\I^{-1}_{\wstar}}
+ 2\sqrt{\tr \left({\I^{-2}_{\wstar}}\right)t}
+ 2\ltwo{\I^{-1}_{\wstar}}t
\right)
\label{eq:vt}
\end{equation}
\end{theorem}

@@

Note that this reduction in variance is essentially optimal.
If we apply Condition 2 to the estimator $\wmle$ instead of $\wmle_i$,
then we are assuming that the variance of $\wmle$ reduces at this same rate.
%By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
%we can show that the variance of the maximum likelihood estimator trained on all of the data is
%\begin{equation}
%\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
%\label{eq:var3}
%\end{equation}
%The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.

@@

%Now we describe our main result.
%First we introduce some notation.
%Let $\proj{\Wowa}$ be the projection operator onto the space $\Wowa$.
For our next results, we need to introduce the following notation.
If $\A$ is an arbitrary space,
then the projection of a point $\x$ onto $\A$ is denoted by
\begin{equation}
\proj{\A}\x = \argmin_{\mathbf{a}\in\A} \ltwo{\mathbf{a}-\x}
\end{equation}
Recall that $\W$ is the vector space spanned by the $\wmle_i$s,
and that $\wowa$ is the maximum likelihood estimator when the parameter space is restricted to $\W$.
Our next result shows that as the number of machines increases,
the distance between $\W$ and $\wstar$ decreases.

\begin{lemma}
Assume Condition 1 holds.
Let $t>0$.
Then we have that with probability at least $1-\exp((d-m)(-t+\ln(t+1)))$,
\begin{equation}
\ltwo{\wstar-\proj\W\wstar}
=
(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
\end{equation}
where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar$.
\end{lemma}

@@

>>>>

%\newcommand{\smin}[1]{s_\text{min}({#1})}
\newcommand{\smin}{s_\text{min}}
\newcommand{\smax}{s_\text{max}}
\newcommand{\qhi}{q_\text{hi}}
\newcommand{\qlo}{q_\text{lo}}
\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Zproj}{Z^{\textit{proj}}}
\newcommand{\Zreopt}{Z^{\textit{owa}}}
\newcommand{\nreopt}{n^{\textit{owa}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\matW}{\hat W}
\newcommand{\matV}{\hat V}
\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Wowa}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\ahat}{\hat\alpha}
\newcommand{\afull}{\ahat^{\textit{full}}}
\newcommand{\wowa}{\hat\w^{owa}}
\newcommand{\wowafull}{\hat\w^{\textit{owa,full}}}
\newcommand{\wowastar}{\hat\w^{\textit{owa,*}}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

In this paper, we propose a \emph{semi-interactive} distributed algorithm called \emph{optimal weighted averaging} (OWA).
Our algorithm performs two rounds of communication,
so it is not subject to the existing regret bounds of non-interactive algorithms.
The amount of information communicated in the second round is small, however,
so our algorithm retains the speed advantages of non-interactive algorithms.

In the next section, we formally describe our problem setting and proposed solution.
In Section 3, we compare our solution to related solutions.
We highlight how the analysis of existing algorithms requires more limiting assumptions than our own,
and show in detail why most existing non-interactive regret bounds do not apply to our problem.
Section 4.1 presents a novel, very general analysis of the averaging estimator.
Section 4.2 shows that two parameters to our algorithm $m$ and $\nreopt$ act as ``knobs'' that let the user easily trade off statistical performance for communication cost.
Section 5 shows experimentally that our algorithm performs well.

\section{Problem Statement and Solution}
%\section{Problem Setting and Our Solution}

@@

%That is, we assume there exists a constant $c$ such that
%\begin{equation}
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
%.
%\end{equation}

\newtheorem*{sgt}{The Sub-Gaussian Tail (SGT) Condition}
\begin{sgt}
Let $\theta$ be an estimator trained on $n$ data points.
Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
\begin{equation}
%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
%\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
\sqrt n \ltwobig{\I^{-1/2}_\wstar(\theta-\E\theta)} < t
,
\end{equation}
where $\I_\wstar$ is the positive definite Fisher information matrix at the parameter vector's true value $\wstar$.
\end{sgt}

@@

%Our condition is more general than the conditions required in previous work.
We require that $\wmle_i$ satisfy the SGT Condition.
This is strictly more general than the conditions in previous work.
\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded (in addition to a number of conditions on the moments of the likelihood).
A bounded parameter space implied that $\wmle_i$ satisfies the SGT Condition because every bounded random variable is automatically sub-Gaussian by definition.

@@

\begin{theorem}
\label{thm:wave}
Assume $\wmle_i$ satisfies the SGT Condition.
Let $t>0$.
Then with probability at least $1-\exp(-t)$,
\begin{equation}
\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
\end{equation}
where
\begin{equation}
v_t =
\sigma^2
\left(
\tr{\I^{-1}_{\wstar}}
+ 2\sqrt{\tr \left({\I^{-2}_{\wstar}}\right)t}
+ 2\ltwo{\I^{-1}_{\wstar}}t
\right)
\label{eq:vt}
\end{equation}
\end{theorem}

@@

Note that this reduction in variance is essentially optimal.
If we assume that $\wmle$ satisfies the SGT Condition,
then we are assuming that the variance of $\wmle$ reduces at the same $O(1/\sqrt{mn})$ rate.
%By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
%we can show that the variance of the maximum likelihood estimator trained on all of the data is
%\begin{equation}
%\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
%\label{eq:var3}
%\end{equation}
%The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.

@@

Unlike the $\wave$ estimator,
$\wowa$ reduces both bias and variance.
The $m$ and $\nreopt$ parameters act as ``knobs'' that let us trade off lower statistical error for higher communication cost.
%In particular, the number of machines $m$ acts as a ``knob'' that lets us
%In this section, we present one main theorem and two supporting lemmas.
%We begin our discussion with the following informal theorem statement.
Before we present the formal analysis,
we present a motivating informal analysis.

\subsubsection{Informal Analysis}
%The following theorem statement ignores constant factors.
%Our main result can be stated with the following informal theorem.
We first state our main result,
then describe its implications.
\newtheorem*{theoreminf}{Theorem 2 (Informal)}
\begin{theoreminf}
With high probability, we have that
\begin{equation}
\ltwo{\wstar-\wowa}
\le
O\left(
    \sqrt{\frac{1}{m\nreopt}}
    +
    \sqrt{\left(1-\frac{m}{d}\right)\ltwo{\wstar-\wave}}
\right)
\label{eq:informal}
\end{equation}
\label{thm:informal}
\end{theoreminf}
The rightmost term captures the error of the optimal parameter vector constrained to $\W$.
That is, if
\begin{equation}
\wowastar = \argmin_{\w\in\W} \int_{\X\times\Y} \left(f(y,\trans\x\w) + \lambda R(\w)\right) d(\x,y)
,
\end{equation}
then the rightmost term is the error of $\ltwo{\wstar-\wowastar}$.
Clearly, as $m \to d$, the space $\W \to \Theta$, so $\wowastar\to\wstar$, and $\ltwo{\wstar-\wowastar}\to 0$.

The leftmost term above captures the error due to using only $\nreopt$ data points in the second round of optimization.
That is, the left most term captures the error $\ltwo{\wowa-\wowastar}$.
As $\nreopt\to n$, this error approaches the error of the oracle estimator trained on all the data $\ltwo{\wstar-\wmle}$.
By increasing $m$ and $\nreopt$ we get an estimator that is closer to the sequential oracle
but at the expense of a higher communication cost.

\subsubsection{Formal Analysis}

In this section we prove a version of Theorem 2 with explicit constant factors.
We begin by introducing the following notation.
The orthogonal projection of a point $\x$ onto the space $\W$ is denoted by
\begin{equation}
\proj{\W}\x = \argmin_{\w\in\W} \ltwo{\w-\x}
.
\end{equation}
Our proof is divided into two lemmas.
The most interesting work happens in the first lemma,
wherein we bound $\ltwo{\wstar-\proj\W\wstar}$.

\begin{lemma}
\label{lem:proj}
Assume $\wmle_i$ satisfies the SGT Condition.
Let $t>0$ and
\begin{equation}
\delta = 1-\exp((d-m)(-t+\ln(t+1)))
.
\label{eq:delta}
\end{equation}
Then we have that with probability at least $\delta$,
\begin{equation}
\ltwo{\wstar-\proj\W\wstar}
=
(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
\end{equation}
where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar$.
\end{lemma}

@@

Our next lemma bounds the error of $\wowa$ in terms of the error of other estimators whose error we know.
%We introduce the following notation.

\begin{lemma}
%Assume that $\wmle_i$ satisfies Condition 1.
%Let $\delta$ be defined as in Equation \ref{eq:delta} from Lemma \ref{lem:proj}.
%Then with probability at least $1-\delta$,
%\begin{equation}
%\ltwo{\wstar-\wowa}
%\le
%\ltwo{\wowastar-\wowa}
%+
%\qlo^{-1} \qhi (
%)
%.
%\end{equation}
Let $\qhi,\qlo : \mathbb{R}^+ \to \mathbb{R}^+$ be monotonically increasing functions such that for all points $\w\in\Theta$,
\begin{align}
F(\wstar) - F(\w) &\ge \qlo \left( \ltwo {\wstar - \w} \right)
,
\\
F(\wstar) - F(\w) &\le \qhi \left( \ltwo {\wstar - \w} \right)
.
\end{align}
Then we have that
\begin{equation}
\ltwo {\wstar-\wowastar}
\le
\qlo^{-1} \left(
    \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
\right)
.
\end{equation}
\end{lemma}

\begin{proof}
By the triangle inequality, we have that
\begin{align}
\ltwo{\wstar-\wowa}
\le
\ltwo{\wstar-\wowastar} + \ltwo{\wowastar-\wowa}
.
\end{align}
We bound the left hand term as follows.
By the definition of $\qhi$ and $\qlo$, we have
\begin{align}
\qlo \left( \ltwo {\wstar-\wowastar} \right)
&\le F(\wstar) - F(\wowastar)
\\
&\le F(\wstar) - F(\proj\W\wstar)
\\
&\le \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
.
\end{align}
%and so
%\begin{equation}
%\ltwo {\wstar-\wowastar}
%\le
%\qlo^{-1} \left(
    %\qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
%\right)
%.
%\end{equation}
%Applying Lemma \ref{lem:proj} gives the stated result.
Applying $\qlo^{-1}$ to both sides gives the stated result.
\end{proof}

We are now ready to state a formal version of the informal Theorem \ref{thm:informal}.

\newtheorem*{theoremformal}{Theorem 2 (Formal)}
\begin{theoremformal}
Assume that both $\wmle_i$ and $\alpha$ satisfy the SGT Condition.
Let $t>0$ and
\begin{equation}
\delta = (1-\exp((d-m)(-t+\ln(t+1))))(1-\exp(-\sigma^2t^2/2))
.
\end{equation}
Then, with probability at least $\delta$,
\begin{equation}
\begin{aligned}
%\ltwo{\wstar-\wowa}
%&\le
%{\wstar-\E\w}
\end{aligned}
\end{equation}
\end{theoremformal}
